---
title: "
isualising_data"
output: html_notebook
---

```{r}
# read the csv file and save as a viriable calle heart_data
setwd("/Users/nadiiakuzmenko/Documents/programming_code/notebooks")

	heart_data<- read.csv("../data/heart_attack_data.csv")
```

```{r}
# preview the data
head(heart_data)
```
```{r}
str(heart_data)
```
```{r}
heart_data$sex <- factor(heart_data$sex)
heart_data$cp <- factor(heart_data$cp)
heart_data$fbs <- factor(heart_data$fbs)
heart_data$restecg <- factor(heart_data$restecg)
heart_data$exang <- factor(heart_data$exang)
heart_data$target <- factor(heart_data$target)
summary(heart_data)
```

```{r}
str(heart_data)
```

```{r}
library(tidyverse)
```
# ggplot(mapping = aes ([MAPPING])) + [GEOM_FUNCTION] ()
A geom is the  name for the specific shape  that  we want  to use to visualize the  data.
To create  a line graph  with ggplot() + geom_line()function.
To create  a scatter  plot, we use ggplot() + geom_point()
To create  a histogram,  we use ggplot() + geom_histogram()
To create  a bar chart  we use ggplot() + geom_bar()
To create  a boxplot,  we use ggplot() + geom_boxplot()



```{r}
heart_data %>%
  # setting variables
  ggplot(aes(x=age, y=chol)) + 
  # ploting scatterplot & adding color & changing size of bubles & fixing issues
  geom_point(aes(color = sex), size=2, na.rm = TRUE) +
#adding title
    ggtitle("Age vs Chol by Gender")

```
```{r}
# adding variable names
heart_data <- heart_data %>% 
  mutate(sex = ifelse(sex == 0, "Female", "Male"))
unique(heart_data$sex)
```

```{r}
# visualizing the relationship between age, chol and heart disease (TARGET)
heart_data %>%
  ggplot(aes(x=age, y=chol, z=target)) +
  geom_point(aes(color= target), na.rm = TRUE) +
    ggtitle("Age Vs Chol and Heart dissease")
```
```{r}
# adding variable names
heart_data <- heart_data %>% 
  mutate(target = ifelse(target == 1, "Disease", "No disease"))
unique(heart_data$target)
```

```{r}
heart_data %>%
  ggplot(aes(x=age)) +
  geom_histogram(fill="pink", color="purple", binwidth = 5, alpha=0.9) +
  ggtitle("Histogram with binwidth 5")
```

